Privacy and Security Benefits of Jailbreaking iOS

### 2013-12-10 final_project/paper.tex
added the tweak. rearranged the paper sections added new stuff

\section{Introduction}

@@

\section{Threats on the App Store}

The App Store is an integral part of the iOS security model. All applications on the App Store must go through an approval process, one of whose purposes is to prevent malicious apps from reaching users. Prior to submitting an app each developer needs to enroll into the paid Apple Developer Program and disclose their identity. This allows Apple to ban developers caught committing infractions from submitting apps to the App Store. While this might discourage developers of malicious apps under the threat of receiving a permanent ban from the App Store, it would be unwise to believe in the benevolence of developers solely due to the existence of a disincentive to misbehaving. Below I outline researchers' findings on circumventing the approval process and examples of malicious apps on the App Store, in order to expose the reliance on the walled garden model as insufficient with regards to ensuring security and privacy.

\subsection{Sneaking malicious code into the AppStore}
Since the approval process does not require that developers submit their source code for a review, it is quite possible to use code obfuscation techniques in order to evade any static or dynamic code analysis tools Apple might be using to uncover threats.

In his paper on iOS privacy\cite{priv}, Nicolas Seriot demonstrates several simple techniques that can be used to deliberately circumvent common types of code analysis likely used by Apple in search of calls to private APIs and system directories. He suggests that his proof-of-concept SpyPhone app\footnote{https://github.com/nst/spyphone/} in fact does not rely on private APIs in the first place, and goes on to express concern that the existence of apps on the App Store collecting personal data is quite likely given the underground demand for and abundance of personal data on iOS devices.

Another group of researchers from Georgia Institute of Technology present\cite{jekyll} their "Jekyll App" which obtained Apple's approval, despite allowing the researchers to remotely launch a variety of attacks on their own tests devices. The paper demonstrates how it is possible to create malicious apps that evade detection by breaking down malicious code into distinct "gadgets" that are not malicious on their own but if  \emph{remotely} supplied the correct execution paths can be assembled intno a malicious routine.

\cite{attacks}, 
\subsection{Threats to Privacy on the AppStore}
Even though Apple tests the applications submitted to the App Store for malicious behavior and security risks, it is even harder to check if apps \emph{misuse} user information. While techniques such as dynamic \cite{dynamic} and static \cite{pios} analysis targeted at capturing data flows within the application have been developed, several controversial cases of user data misuse and user profiling have gained attention\footnote{http://mclov.in/2012/02/08/path-uploads-your-entire-address-book-to-their-servers.html}. The problem of course is somewhat intractable, because trusting data to a third party comes with no guarantees about how the data will be used.

Nevertheless, it is important to make sure that users are given the power to \emph{decide} whether to give their data in the first place. This privacy enhancement feature can be a great complement to pre-existent code approval policies and jailbreaking has allowed for several such data flow management tools. However it wasn't until iOS7 before users were able to control microphone use or limit ad tracking\footnote{http://blogs.wsj.com/digits/2013/09/18/how-to-use-apples-new-ios-7-privacy-controls/} and a dedicated Privacy Settings configuration pane providing consolidated control over the permissions of individual apps has been available only as of iOS6\footnote{https://developer.apple.com/library/ios/releasenotes/General/RN-iOSSDK-6\_0/}.

\section{Action Items}
\subsection{Jailbreaking for better privacy and security}

@@

\section{Conclusion}

>>>>

\section{Introduction}
\tab The App Store is an integral part of the iOS security model. All applications on the App Store must go through an approval process, one of whose purposes is to prevent malicious apps from reaching users. This process involves quality control, static and manual analysis of the behavior and content of apps. In addition the use of private APIs, which typically would allow an app to provide unapproved or malicious functionality is screened against. Apps on the iOS platform are assigned a unique identifier and a folder whose contents it can modify, while the rest of the filesystem is out of reach. Individual apps can communicate with other apps in a very limited fashion by means of URL Handlers. Access to Apple services such as Push Notifications, Game Center and Passbook is authenticated through certificates issued by Apple. Moreover, every application on the App Store is code signed, and code that hasn't been signed by Apple cannot run on the device. All these measures ensure that apps run in a sandboxed environment which ensures that the system and the user data on the device are protected in case an app attempts malicious activity.

Jailbreaking is the process of modifying iOS through what is most often a security exploit in order to allow code that hasn't been signed by Apple to run on the device, in addition to enabling root access to the filesystem for any app. Hence jailbreaking allows for downloading apps from sources exempt of Apple's stringent approval procedures, which typically enhance the functionality of the OS and its stock apps and allow for unapproved use of private API that provide greater functionality and freedom to developers.

 Below I outline researchers' findings on circumventing the approval process and examples of malicious apps on the App Store, in order to expose the reliance on the walled garden model as insufficient with regards to ensuring security and privacy.

@@

\section{Sneaking malicious code into the AppStore}
\tab Prior to submitting an app each developer needs to enroll into the paid Apple Developer Program and disclose their identity. This allows Apple to ban developers caught committing infractions from submitting apps to the App Store. While this might discourage developers of malicious apps under the threat of receiving a permanent ban from the App Store, it would be unwise to believe in the benevolence of developers solely due to the existence of a disincentive to misbehaving.

In reality, since the approval process does not require that developers submit their source code for a review, it is quite possible to use code obfuscation techniques in order to evade any static or dynamic code analysis tools Apple might be using to uncover threats.  In his paper on iOS privacy\cite{priv}, Nicolas Seriot demonstrates several simple techniques that can be used to deliberately circumvent common types of code analysis likely used by Apple in search of calls to private APIs and system directories. He suggests that his proof-of-concept SpyPhone app\footnote{https://github.com/nst/spyphone/} in fact does not rely on private APIs in the first place, and goes on to express concern that the existence of apps on the App Store collecting personal data is quite likely given the underground demand for and abundance of personal data on iOS devices.

Another group of researchers from Georgia Institute of Technology present\cite{jekyll} their "Jekyll App" which obtained Apple's approval, despite allowing the researchers to remotely launch a variety of attacks on their own tests devices. The paper demonstrates in practice how it is possible to create malicious apps that evade detection by breaking down malicious code into distinct subroutines termed "gadgets" which are not malicious on their own but if  \emph{remotely} supplied the correct execution paths can be assembled into a malicious routine. Breaking down code with malicious effects into innocuous looking subroutines practically renders static analysis less useful as the malicious code no longer appears explicitly. Moreover, since the malicious routines are enabled remotely, even the effectiveness of manual testing and dynamic analysis is diminished.

\cite{attacks}, 
\section{Threats to Privacy in the AppStore}
Even though researchers have demonstrated techniques such as dynamic \cite{dynamic} and static \cite{pios} analysis targeted at capturing data flows within the application have been developed, several controversial cases of user data misuse and user profiling have gained attention\footnote{http://mclov.in/2012/02/08/path-uploads-your-entire-address-book-to-their-servers.html}.

A recent case of privacy violations emerged in November 2013, as the developer Kyle Richter discovered that a popular game app QuizUp transmitted sensitive information about a players Facebook friends, Address Book contacts and others, in a plain text format over to the company servers, presumably in an effort to match the user with other players.\footnote{http://kylerichter.com/our-responsibility-as-developers/} Despite the fact that thanks to the privacy enhancements in iOS, the app had to request users' permission to collect the data, the users are still not able to fine tune the permissions granted, for example by disallowing access to contacts' email addresses or real names, which wouldn't be of use to a game matching algorithm.

While AppStore apps can pose threats to privacy, iOS itself collects large amounts of data from normal usage.
In 2011, a controversy arose due to iOS maintaining a database of WiFi hotspots and cell towers around the users location, in order to improve its geo-location services beyond complete reliance on the slower GPS systems which would allow for quickly determining the user's current location when using Maps.\footnote{http://www.networkworld.com/community/blog/apple-officially-responds-ios-consolidateddb-} Even though, Apple assured that despite this location data being sent to Apple servers, it is completely anonymous the data would remain on the device itself in a file named consolidated.db which can then be easily retrieved.\footnote{https://www.sans.org/reading-room/whitepapers/forensics/forensic-analysis-ios-devices-34092}.

A new feature in iOS7, called Frequent Locations relies on a similar location data storage, which contains the users most frequently visited locations. The idea behind the feature is to enable Next Destination, a user-end feature that lets Apple provide relevant commute information such as travel time to work.\footnote{http://www.idownloadblog.com/2013/10/20/how-to-use-frequent-locations-and-next-destination-features-in-ios-7/} Despite users being able to disable those features, they are enabled by default and the data collection is not apparent.

\section{Threats of Jailbreaking}

you can crack the pin code of someone by jailbreaking an un jailbroken phone first and then changing the pin
you can use the espionage tools in \cite{espionage} unless you have stuff fixed.
default root password is very well known.
can't update if u want to keep the jailbreak.
installing malicious apps from evil repositories.

\section{Action Items}
\subsection{Jailbreaking for better privacy}

Before address book permissions were first added to iOS 6, the well publicized case of the social networking app Path drew attention to the issue of privacy in the AppStore. It was discovered that Path collected personal information from the user's address book and sent it to company servers without the users consent and it did not list this activity in the privacy agreement. The creators of Path have since removed the behavior and settled the ensuing FTC charges. \footnote{http://www.ftc.gov/news-events/press-releases/2013/02/path-social-networking-app-settles-ftc-charges-it-deceived} In response to this, a tweak for jailbroken iOS devices named ContactPrivacy appeared. It allowed the user to control access to the address book.

@@

\section{Conclusion}
The problem of course is somewhat intractable, because trusting data to a third party comes with no guarantees about how the data will be used. Nevertheless, it is important to make sure that users are given the power to \emph{decide} whether to give their data in the first place. This privacy enhancement feature can be a great complement to pre-existent code approval policies and jailbreaking has allowed for several such data flow management tools. However it wasn't until iOS7 before users were able to control microphone use or limit ad tracking\footnote{http://blogs.wsj.com/digits/2013/09/18/how-to-use-apples-new-ios-7-privacy-controls/} and a dedicated Privacy Settings configuration pane providing consolidated control over the permissions of individual apps has been available only as of iOS6\footnote{https://developer.apple.com/library/ios/releasenotes/General/RN-iOSSDK-6\_0/}.
